Self Referencing Sequences

### 2014-10-28 draft.tex
Edited equivalence

In this section, we consider a slightly weaker definition of equivalence. Consider, for example, sequences generated by $C = \{1,2,3\}$. Theorem~\ref{independentsequences} shows that the sequence starting with 2
\begin{align*}
2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
and the sequence starting with 3
\begin{align*}
3,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
are independent in the sense that neither sequence is exactly contained in the other. However, if we delete the first two numbers of the sequence starting with 2, and the first number of the sequence starting with 3, we get 
\begin{align*}
3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
in both cases. Lets look at the state of the algorithm that generates both sequences. After the first block, the execution of the algorithm of $s(C, 2)$ right after placing the $3$'s (in the 4th and 5th position), the algorithm is in state corresponding to:
\begin{itemize}
\item block size $3$
\item no elements on the current block
\item starting number is $1$
\end{itemize}.   
If we look at the execution of the algorithm that generates $s(C, 3)$ after placing the first three $3$'s, the algorithm is in state corresponding to:
\begin{itemize}
\item block size $3$
\item no elements on the current block
\item starting number is $1$
\end{itemize}
Therefore, since the states are the same, the same sequence will be generated after the 4th position of $s(C, 2)$ and after the $3rd$ position of $s(C, 3)$ will be the same. By inspection, we can see we can include $2$ of the 3's.

Thus, since both sequences are essentially the same if certain prefixes are removed, we call these sequences \emph{equivalent}. To formalize, this, we say that two sequences $\{a_i\}$ and $\{b_i\}$ are equivalent if there exists some integers $n,k \geq 0$ such that
\begin{align*}
a_{n+i} = b_{k+i}
\end{align*}

@@

This conjecture is partly motivated by the fact that the property holds for $\{1,2,3\}$. For generating sets with larger numbers, we have been unable to verify or refute the conjecture. However, in Section~\ref{Density}, we show that the starting number of a sequence does not affect the densities of each number, which provides some evidence for this conjecture.

We now introduce a notion which will help us categorize sequences as equivalent. 
\begin{definition} Consider a sequence $x$. We define the read distance for an element $x_i$ as the number of terms in between it and the beginning of the $i$th block of numbers. We denote this as $Rd(x_i)$. 
\end{definition}
The read distance can be any nonnegative integer. For example if we consider the sequence $a_i$ generated by $\{1,2,3\}$ beginning with a 2  
\begin{align*}
2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
the read distance for the $a_1$ (the first two) is $0$ as the first block of numbers starts at the first number, the read distance for $a_2$ (the second two) is $1$ as the second block of numbers starts at the third number. We can continue to calculate the read distances for this sequence and see that they form a monotonically increasing sequence $0,1,2,4,6,6,6,6,7,8,9,\ldots$. From this we might suspect that the read distance for any self referencing sequence will also be monotonically increasing. We can in fact show a much stronger result. 
\begin{lemma} 
\label{readdistformula}
Let $a_i$ be any self referencing sequence generated by some set of positive integers. Then the read distance for the element $a_i$ for $i>1$ is 
\begin{equation*}
\sum_{k=1}^{i-1}( a_k - 1)
\end{equation*} 
\end{lemma}
\begin{proof}
We will prove this using induction. We know that for $i=1$ the read distance is $0$, as the first block always starts at the first number. For the base case $i=2$ we must show that the read distance is $a_1-1$. Well we know that the first block will be of length $a_1$ since the sequence is self referencing, therefore the second block will begin at \emph{position} $a_1$. The number of elements between position $a_1$ and $2$ is exactly $a_1 -1$.

Now if we assume that our result holds for $i=l$ we would like to show that it holds for $i=l+1$. We know that the distance between the $l$th block and the number at position $l$ is 
\begin{equation*}
\sum_{k=1}^{l-1}( a_k - 1).
\end{equation*} 
This $l$th block has length $a_l$ and so the $l+1$th block will begin $a_l$ positions to the right of the beginning of the $l$th block. Also note that $a_{l+1}$ is one position to the right of $a_l$. Therefore the beginning of the $l+1$th block is 
\begin{equation*}
a_l-1 + \sum_{k=1}^{l-1} (a_k-1) = \sum_{k=1}^{l} (a_k-1)
\end{equation*} 
positions to the right of $a_{l+1}$. But this is exactly the reading distance for $a_{l+1}$ and so we have our result.
\end{proof}
Note that because $a_i$ is generated by positive integers we always have $a_k-1\geq 0$ and the read distance is monotonically increasing. Although this shows us that read distance never decreases it does not tell us how quickly it increases, our next result gives a bound on the number of consecutive times the read distance can stay constant.

\begin{lemma} Consider a self referencing sequence $x$ generated by $\{a_1, a_2, \ldots, a_n\}$ with $a = \max{ a_1, a_2, \ldots, a_n}$. Suppose the read distance for $x_i$ is $d$, then the read distance for $x_{i+a}$ is at least $d+1$. 
\end{lemma}
\begin{proof}
Assume otherwise. By our Lemma ~\ref{readdistformula} we know that the read distance for $x_{i+a}$ is at least $d$. By our assumption it must be exactly $d$. Lemma ~\ref{readdistformula} also tells us that $x_{i+k}=1$ for $k=0,1,2,\ldots, a_n$. But this means that we have a block of length $a+1$, which is a contradiction of the self referencing property of the sequence. Therefore the read distance for $x_{i}$ is at least $d+1$. 
\end{proof}

>>>>

In this section, we consider a slightly weaker definition of equivalence. Consider, for example, sequences generated by $G = \{1,2,3\}$. Theorem~\ref{independentsequences} shows that the sequence starting with 2
\begin{align*}
2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
and the sequence starting with 3
\begin{align*}
3,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
are independent in the sense that neither sequence is exactly contained in the other. However, if we delete the first two numbers of the sequence starting with 2, and the first number of the sequence starting with 3, we get 
\begin{align*}
3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
in both cases. Lets look at the state of the algorithm that generates both sequences. After the first block, the execution of the algorithm of $s(G, 2)$ right after placing the first $3$'s (in the 3rd and 4th position), the algorithm is in state corresponding to:
\begin{itemize}
\item block size $3$
\item no elements in the current block
\item starting number is $1$
\end{itemize}.   
If we look at the execution of the algorithm that generates $s(G, 3)$ after placing the first three $3$'s, the algorithm is in state corresponding to:
\begin{itemize}
\item block size $3$
\item no elements in the current block
\item starting number is $1$
\end{itemize}
Therefore, since the states are the same, the sequence generated after the 4th position of $s(G, 2)$ and after the $3rd$ position of $s(G, 3)$ will be the same. By inspection, we see we can also include two of the 3's.

Thus, since both sequences are essentially the same if a prefix is removed from each of them, we call these sequences \emph{equivalent}. To formalize, this, we say that two sequences $\{a_i\}$ and $\{b_i\}$ are equivalent if there exists some integers $n,k \geq 0$ such that
\begin{align*}
a_{n+i} = b_{k+i}
\end{align*}

@@

This conjecture is partly motivated by the fact that the property holds for $\{1,2,3\}$. For generating sets with larger numbers, we have been unable to verify or refute the conjecture. However, in Section~\ref{density}, we show that for some generating sets the starting number of a sequence does not affect the densities of each number, which provides some evidence for this conjecture.

We now introduce a notion which will help us categorize sequences as equivalent. 
\begin{definition} Consider a sequence $a$. We define the read distance for an element $a_i$ as the number of terms in between it and the beginning of the $i$th block of numbers.
\end{definition}
The read distance can be any nonnegative integer. For example if we consider the sequence $a_i$ generated by $\{1,2,3\}$ beginning with a 2  
\begin{align*}
2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
the read distance for $a_1$ (the first two) is $0$ as the first block of numbers starts at the first number, the read distance for $a_2$ (the second two) is $1$ as the second block of numbers starts at the third number. We can continue to calculate the read distances for this sequence and see that they form a monotonically increasing sequence $0,1,2,4,6,6,6,6,7,8,9,\ldots$. From this we might suspect that the read distance for any self referencing sequence will also be monotonically increasing. We can in fact compute the read distance exactly. 
\begin{lemma} 
\label{readdistformula}
Let $a_i$ be any self referencing sequence generated by some set of positive integers. Then the read distance for the element $a_i$, for $i>1$, is 
\begin{equation*}
\sum_{k=1}^{i-1}( a_k - 1)
\end{equation*} 
\end{lemma}
\begin{proof}
We will prove this using induction. We know that for $i=1$ the read distance is $0$, as the first block always starts at the first number. When $i=2$ we must show that the read distance is $a_1-1$. The first block will be of length $a_1$ since the sequence is self referencing, therefore the second block will begin at position $a_1$. The number of elements between position $a_1$ and $2$ is exactly $a_1 -1$.

Now if we assume that our result holds for $i=l$ we would like to show that it holds for $i=l+1$. We know that the distance between the start of the $l$th block and position $l$ is 
\begin{equation*}
\sum_{k=1}^{l-1}( a_k - 1).
\end{equation*} 
This $l$th block has length $a_l$ and so the $l+1$th block will begin $a_l$ positions to the right of the beginning of the $l$th block. Also note that $a_{l+1}$ is one position to the right of $a_l$. Therefore the beginning of the $(l+1)$th block is 
\begin{equation*}
a_l-1 + \sum_{k=1}^{l-1} (a_k-1) = \sum_{k=1}^{l} (a_k-1)
\end{equation*} 
positions to the right of $a_{l+1}$. But this is exactly the reading distance for $a_{l+1}$ and so we have our result.
\end{proof}
Note that because $a_i$ is generated by positive integers we always have $a_k-1\geq 0$ and so the read distance is monotonically increasing. Although this shows us that read distance never decreases it does not tell us how quickly it increases, our next result gives a simple bound on the number of consecutive times the read distance can stay constant.

\begin{lemma} Consider a self referencing sequence $x$ generated by $\{a_1, a_2, \ldots, a_n\}$ with $a = \max(a_1, a_2, \ldots, a_n)$. Suppose the read distance for $x_i$ is $d$, then the read distance for $x_{i+a}$ is at least $d+1$. 
\end{lemma}
\begin{proof}
Assume otherwise. By Lemma ~\ref{readdistformula} we know that the read distance for $x_{i+a}$ is at least $d$. Thus it must be exactly $d$. Lemma ~\ref{readdistformula} also tells us that if this is the case then $x_{i+k}=1$ for $k=0,1,2,\ldots, a_n$. But this means that we have a block of length $a+1$, which is a contradiction of the self referencing property of the sequence. Therefore the read distance for $x_{i}$ is at least $d+1$. 
\end{proof}
